
\section[Methods]{Methods}
\label{sec:methods}
\addcontentsline{toc}{section}{\thesection. Methods}

Some details can be found on wikipedia website at
\begin{center}
\url{https://en.wikipedia.org/wiki/Eigendecomposition_of_a_matrix}
\end{center}
for generalized eigenvalues, and at
\begin{center}
\url{https://en.wikipedia.org/wiki/Schur_decomposition}
\end{center}
about QZ decomposition or generalized Schur form.
The LAPACK~\citep{laug} also provides functions to
solve these problems.


\subsection[Generalized Eigenvalues for Pair Matrices]{Generalized Eigenvalues for Pair Matrices}
\label{sec:generalized_eigenvalues_pair}
\addcontentsline{toc}{subsection}{\thesubsection. Generalized Eigenvalues for Pair Matrices}

Suppose $\bA$ and $\bB$ are two $N\times N$ non-symmetric matrices
which can be both in real or in complex. The goal is to find right generalized
eigen vectors $\bv$ such that $\bA \bv = \lambda \bB \bv$, or left generalized
eigen vectors $\bu$ such
that $\bu^{H} \bA = \lambda \bu^{H} \bB$ where $\bu^{H}$ is the
conjugate-transpose of $\bu$.
Also, $\lambda$ is called generalized eigenvalues of $\bA$ and $\bB$
which obeys $\mbox{det}(\bA - \lambda\bB) = 0$.
Note that $\lambda$, $\bu$, and $\bv$ may be complex even $\bA$ and $\bB$
are in real.

Suppose $\bB$ is an identity matrix $\bI$, then the problem reduces to
traditional eigenvalue problem. i.e. This is a special case.


\subsection[QZ Decomposition for Pair Matrices]{QZ Decomposition for Pair Matrices}
\label{sec:qz_decomposition_pair}
\addcontentsline{toc}{subsection}{\thesubsection. QZ Decomposition for Pair Matrices}

Suppose $\bA$ and $\bB$ are two $N\times N$ non-symmetric matrices
which can be both in real or in complex.
The QZ decomposition factorizes both matrices as
\begin{itemize}
\item
$\bA = \bQ\bS\bZ^\top$ and $\bB = \bQ\bT\bZ^\top$
if $\bA$ and $\bB$ are real, or
\item
$\bA = \bQ\bS\bZ^{H}$ and $\bB = \bQ\bT\bZ^{H}$
if $\bA$ and $\bB$ are complex
\end{itemize}
where $\bQ$ and $\bZ$ are unitary and $\bS$ and $\bT$ are upper triangular.
The unitary means $\bX\bX^{H} = \bI$ if $\bX$ is complex or
$\bX\bX^\top = \bI$ if $\bX$ is real where $\bI$ is the identity matrix.

The QZ decomposition is also called generalized Schur decomposition
where $\bS$ and $\bT$ are the Schur form of $\bA$ and $\bB$.
The generalized eigenvalues $\lambda$ that solve the generalized eigenvalue
problem $\bA \bx = \lambda \bB \bx$ where $\bx$ is an unknown nonzero vector
and $\lambda_i = \bS_{ii} / \bT_{ii}$.

Suppose $\bB$ is an identity matrix $\bI$, then the problem reduces to
fine $\bQ$ such that $\bA = \bQ\bS\bQ^{-1}$ for real $\bA$ or
$\bA = \bQ\bS\bQ^{H}$ for complex $\bA$.
i.e. This is a special case.


